---
title: "Homework 3 zz2901"
author: "Troy Zhou"
output: github_document
---

```{r load_package, echo = FALSE, message=FALSE}
library(tidyverse)
options(tibble.print_max = 5)
```

## Problem 2

Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices measure “activity counts” in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The data can be downloaded here. **In this spreadsheet, variables `activity.*` are the activity counts for each minute of a 24-hour day starting at midnight.**

* Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel_data <- read.csv("data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(real_day_id=day_id-(week-1)*7,
    day = recode(real_day_id,`1`="Monday",`2`="Tuesday",`3`="Wednesday",
                 `4`="Thursday",`5`="Friday",`6`="Saturday",`7`="Sunday"),
        weekday_vs_weekend = if_else(day=="Saturday"|day=="Sunday",1,2,missing=99),
         weekday_vs_weekend = recode(weekday_vs_weekend,`1`="Weekend",`2`="Weekday")) %>% 
  select(1,2,3,weekday_vs_weekend,everything(),-real_day_id)

accel_data[1:5,1:5]
```

<span style="color:#36648B;font-weight:700">

* Description of `accel_data`

  * `accel_data` was imported from the `csv` file `accel_data.csv` with a dimension of ``r nrow(accel_data)`` observations and ``r ncol(accel_data)-1`` variables. After adding the `weekday_vs_weekend` variable, there are ``r ncol(accel_data)`` variables. 
  
  * the `day` variable might have been labeled wrongly. It was corrected based on `day_id`.
  
  * `chr` variables `day` and `weekday_vs_weekend`, and `num` variables `week`, `day_id`, and `activity_*` are present in the dataset after tidying. 

</span>

* Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_aggregate <- accel_data %>% 
  pivot_longer(5:ncol(accel_data),values_to = "total_activity") %>% 
  group_by(day_id) %>% 
  mutate(total_activity=sum(total_activity)) %>% 
  distinct(day_id,.keep_all=TRUE) %>% 
  select(everything(),-name)

table_prob1 <- accel_aggregate%>% 
  knitr::kable(digits = 3)
```

<span style="color:#36648B;font-weight:700">

* This table shows the aggregated `activity.*` for each `day`.

  * Table: `r table_prob1` The trends are really hard to see from a table. But we do notice there was almost always one day a week that he had significantly lower total activity compared to the rest of the week.

</span>

* Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_aggregate %>%
  ggplot(aes(x = day_id, y = total_activity)) + 
    geom_point(aes( color = day)) + geom_smooth(alpha=0.3)+
    geom_line(aes(color=day),alpha=.3)+
    theme(legend.position = "bottom")+
    labs(
    title = "Q1. 24-hour activity time courses for each day",
    x = "Day of Data Collection",
    y = "24-hour Total Activity")
```

<span style="color:#36648B;font-weight:700">

* description and conclusion:

  * From the scatterplot and trend lind created by `geom_point()` and `geom_smooth()`, we can observe that the subject's total activity fluctuates greatly throughout the 35 days and also days fo each week. He seems to have a period with higher activity after one with lower activity and vice versa from the graph. The trends of activity for day of the week follow the general trends as shown by the lines created from `geom_line()`.
  
</span>

## Problem 3

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

* Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

<span style="color:#36648B;font-weight:700">

</span>

* Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

<span style="color:#36648B;font-weight:700">

</span>

* Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

<span style="color:#36648B;font-weight:700">

</span>


